### 2016-07-24 Sup1.Rmd
図1を一新。図2 (機構のパス図) は削除。SPDはSupへ。 もう一息。3295 words

# Supplemental Information S1

Information of the light sources and measuring instruments.

| Light                           | Model code        | Supplier                        | Measuring instrument              |
|:---------                       |:-----------       |:---------                       |:---------------------             |
| Fluorescent lamp (FL)           | FPL55EX-N         | IWASAKI ELECTRIC CO., LTD.      | HR2000 (Ocean Optics Inc.)        |
| Metalhalide lmap (MH)           | M220FCELS-W/BUD   | IWASAKI ELECTRIC CO., LTD.      | HR2000 (Ocean Optics Inc.)        |
| High-pressure sodium lamp (HPS) | NH220FLS          | IWASAKI ELECTRIC CO., LTD.      | HR2000 (Ocean Optics Inc.)        |
| Blue LED (B)                    | HBL3-3S55-LE      | Tricon Co.                      | MS-720 (EKO Instruments Co. Ltd.) |
| Red LED (R)                     | SRK3-3A80-LE      | Tricon Co.                      | MS-720 (EKO Instruments Co. Ltd.) |
| Far-red LED (FR)                | L735-36AU         | Epitex Inc.                     | MS-720 (EKO Instruments Co. Ltd.) |
| White LED (W)                   | NSPW310DS         | Nichia Chemical Industries Ltd. | MS-720 (EKO Instruments Co. Ltd.) |

The SPD of sunlight was adapted from JIS C8904-3.
The SPD of the shaded sunlight was calculated from the SPD of sunlight and the transmittance spectrum of a cucumber leaf (Murakami et al. 2016).

```{r FigSPD, eval = evalFig, fig.cap = CapSPD}
SPDs
```