### 2016-08-04 NotUsed/Sup1.Rmd
英文構成を依頼 (Edanz)。3089 words

# Supplemental Information S1

Information of the spectral distributions of artificial assimilation light and sunlight.

<!-- Table S1.  Description of the light sources and the instruments used for the measurements of the spectral photon flux density distributions.  -->

<!-- | Light source (abbreviation)     | Model code                    | Measuring instrument              | -->
<!-- |:--------------------------------|:------------------------------|:----------------------------------| -->
<!-- | Fluorescent lamp (FL)           | FPL55EX-N<sup>\*1</sup>       | HR2000 (Ocean Optics Inc.)        | -->
<!-- | Metal halide lamp (MH)          | M220FCELS-W/BUD<sup>\*1</sup> | HR2000 (Ocean Optics Inc.)        | -->
<!-- | High-pressure sodium lamp (HPS) | NH220FLS<sup>\*1</sup>        | HR2000 (Ocean Optics Inc.)        | -->
<!-- | Blue LED (B)                    | HBL3-3S55-LE<sup>\*2</sup>    | MS-720 (EKO Instruments Co. Ltd.) | -->
<!-- | Red LED (R)                     | SRK3-3A80-LE<sup>\*2</sup>    | MS-720 (EKO Instruments Co. Ltd.) | -->
<!-- | Far-red LED (FR)                | L735-36AU<sup>\*3</sup>       | MS-720 (EKO Instruments Co. Ltd.) | -->
<!-- | White LED (W)                   | NSPW310DS<sup>\*4</sup>       | MS-720 (EKO Instruments Co. Ltd.) | -->

<!-- ^\*1^IWASAKI ELECTRIC CO., LTD., Tokyo, Japan; ^\*2^Tricon Co., Shimane, Japan; ^\*3^Epitex Inc., Kyoto, Japan; ^\*4^Nichia Chemical Industries Ltd., Tokushima, Japan. -->

<!-- | Light (abbreviation)            | Model code        | Supplier                        | Measuring instrument              | -->
<!-- |:--------------------------------|:------------------|:--------------------------------|:----------------------------------| -->
<!-- | Fluorescent lamp (FL)           | FPL55EX-N         | IWASAKI ELECTRIC CO., LTD.      | HR2000 (Ocean Optics Inc.)        | -->
<!-- | Metal halide lmap (MH)          | M220FCELS-W/BUD   | IWASAKI ELECTRIC CO., LTD.      | HR2000 (Ocean Optics Inc.)        | -->
<!-- | High-pressure sodium lamp (HPS) | NH220FLS          | IWASAKI ELECTRIC CO., LTD.      | HR2000 (Ocean Optics Inc.)        | -->
<!-- | Blue LED (B)                    | HBL3-3S55-LE      | Tricon Co.                      | MS-720 (EKO Instruments Co. Ltd.) | -->
<!-- | Red LED (R)                     | SRK3-3A80-LE      | Tricon Co.                      | MS-720 (EKO Instruments Co. Ltd.) | -->
<!-- | Far-red LED (FR)                | L735-36AU         | Epitex Inc.                     | MS-720 (EKO Instruments Co. Ltd.) | -->
<!-- | White LED (W)                   | NSPW310DS         | Nichia Chemical Industries Ltd. | MS-720 (EKO Instruments Co. Ltd.) | -->

#### Supplementary Reference
JIS C 8904-3. 2011. [Photovoltaic devices-Part 3: Measurement principles for photovoltaic(PV) solar devices with reference spectral irradiance data](http://www.jisc.go.jp/app/pager?id=2769864).