### 2016-08-16 Header.Rmd
3172 words. KF, RMのコメントを反映。

---
title: "A basis for selecting light spectral distribution for evaluating photosynthetic rates of leaves grown under different light spectral distributions"
author: "Keach Murakami^1\\*^, Ryo Matsuda^1^, Kazuhiro Fujiwara^1^"
csl: ECB.csl
bibliography: ~/GitHub/BeeLabR/BibTex/With.bib
header-includes:
  - \setmainfont{Times New Roman}
  # - \setmainfont{Helvetica Neue}
  - \usepackage{zxjatype} # 日本語を使うためのパッケージ
  - \setjamainfont{Hiragino Kaku Gothic Pro} # 日本語フォントの指定
  - \usepackage{lscape} # 図を回転するためのパッケージ
  - \newcommand{\blandscape}{\begin{landscape}} # 図回転用のbegin command
  - \newcommand{\elandscape}{\end{landscape}} # 図回転用のend command
  - \usepackage{setspace} # 行間の調節ためのパッケージ
  - \newcommand{\bdoublespace}{\begin{spacing}{1.5}} # 行間を広げるbegin command
  - \newcommand{\edoublespace}{\end{spacing}{1.5}} # 行間を広げるbegin command
  - \usepackage{lineno} # 行番号を追加するためのパッケージ
  - \linenumbers # 行番号を追加する

output:
#   pdf_document:
#     latex_engine: xelatex
#     fig_caption: TRUE
# fontsize: 20pt
# geometry: margin=1.2in
  html_document:
    toc: TRUE
    depth: 3
    theme: sandstone
    self_contained: TRUE
    fig_caption: TRUE
  word_document:
    toc: FALSE
    fig_caption: TRUE
---

```{r, echo = F, message = F}
rm(list = ls())

evalFig <- T
base_size <- 20
Annotate_size <- 5

source("./caption.R")
if(evalFig){
  devtools::source_url("https://raw.githubusercontent.com/KeachMurakami/Sources/master/Startup_silent.R")
}

knitr::opts_chunk$set(cache = F, echo = FALSE, fig.width = 7, fig.height = 7, dpi = 300,
                      message = FALSE, warning = FALSE, eval = T)
               # out.width = '500pixel', out.height = '500pixel', dpi = 100,
```

Affiliations: ^1^Graduate School of Agricultural and Life Sciences, The University of Tokyo

Address: ^1^Yayoi, Bunkyo, Tokyo, 113--8657, Japan

e-mail: keach.murakami@gmail.com

**Abstract**

  The spectral photon flux density distributions of light during growth and for measurements do not only directly affect the net photosynthetic rate (*P*~n~).
  Moreover, 
  This paper summarizes a plausible mechanism of the interaction, some situations in which the interaction should be considered, and recommendations for selecting appropriate measuring light to evaluate photosynthesis.
  The *P*~n~ should be measured under *in situ* conditions, depending on the purpose of the study.
  For studies focusing on the evaluation of plant growth after measurements, *P*~n~ should be measured under the SPD of light that the plant will be subjected to.
  In retrospective studies aiming to elucidate the causes of differences among experimental groups, measurements should be made under the SPDs of the lights the plants were grown under.
  In descriptive studies, *P*~n~ should be evaluated under several SPDs of measuring light to understand the general photosynthetic characteristics of the leaves.
  The *P*~n~ under a single spectral distribution of measuring light is only one aspect of the photosynthetic characteristics of a leaf.
  The obtained results must be discussed in relation to the SPD of measuring light so as not to make biased evaluations.

------

**Abbreviations**
BR-light: blue and red LED light,
Chl: chlorophyll,
ETR: electron transport rate,
FR: far-red,
LED: light-emitting diode,
*P*~n~: net photosynthetic rate,
PFD: photon flux density,
PPFD: photosynthetic photon flux density,
PSI: photosystem I,
PSII: photosystem II,

------

```{r, child = "./Main.Rmd"}
```

# References