### 2016-08-18 Header.Rmd
3363 words おおづめ

---
title: "A basis for selecting light spectral distribution for evaluating photosynthetic rates of leaves grown under different light spectral distributions"
author: "Keach Murakami^1\\*^, Ryo Matsuda^1^, Kazuhiro Fujiwara^1^"
csl: ECB.csl
bibliography: ~/GitHub/BeeLabR/BibTex/With.bib
header-includes:
  - \setmainfont{Times New Roman}
  # - \setmainfont{Helvetica Neue}
  - \usepackage{zxjatype} # 日本語を使うためのパッケージ
  - \setjamainfont{Hiragino Kaku Gothic Pro} # 日本語フォントの指定
  - \usepackage{lscape} # 図を回転するためのパッケージ
  - \newcommand{\blandscape}{\begin{landscape}} # 図回転用のbegin command
  - \newcommand{\elandscape}{\end{landscape}} # 図回転用のend command
  - \usepackage{setspace} # 行間の調節ためのパッケージ
  - \newcommand{\bdoublespace}{\begin{spacing}{1.5}} # 行間を広げるbegin command
  - \newcommand{\edoublespace}{\end{spacing}{1.5}} # 行間を広げるbegin command
  - \usepackage{lineno} # 行番号を追加するためのパッケージ
  - \linenumbers # 行番号を追加する

output:
#   pdf_document:
#     latex_engine: xelatex
#     fig_caption: TRUE
# fontsize: 20pt
# geometry: margin=1.2in
  html_document:
    toc: TRUE
    depth: 3
    theme: sandstone
    self_contained: TRUE
    fig_caption: TRUE
  word_document:
    toc: FALSE
    fig_caption: TRUE
---

```{r, echo = F, message = F}
rm(list = ls())

evalFig <- T
base_size <- 20
Annotate_size <- 5

source("./caption.R")
if(evalFig){
  devtools::source_url("https://raw.githubusercontent.com/KeachMurakami/Sources/master/Startup_silent.R")
}

knitr::opts_chunk$set(cache = F, echo = FALSE, fig.width = 7, fig.height = 7, dpi = 300,
                      message = FALSE, warning = FALSE, eval = T)
               # out.width = '500pixel', out.height = '500pixel', dpi = 100,
```

Affiliations: ^1^Graduate School of Agricultural and Life Sciences, The University of Tokyo

Address: ^1^Yayoi, Bunkyo, Tokyo, 113--8657, Japan

e-mail: keach.murakami@gmail.com

**Abstract**
  Relative spectral distributions of light during growth and for measurements do not only directly affect the net photosynthetic rate (*P*~n~), but also indirectly affect it through the interaction between the distributions of the growth and the measuring light.
  This paper summarizes a plausible mechanism of the interaction, some situations in which the interaction should be considered, and recommendations for selecting appropriate measuring light to evaluate photosynthesis.
  In agricultural and horticultural studies, the *P*~n~ should be measured under *in situ* conditions, depending on the purpose of the study.
  For prospective studies focusing on the evaluation of plant growth after measurements, *P*~n~ should be measured under the spectral distribution of light to which the plant will be subjected.
  In retrospective studies aiming to elucidate the causes of differences in growth brought about by the different growth conditions, measurements should be made under the spectral distributions of the lights under which the plants were grown.
  The *P*~n~ under a single spectral distribution of measuring light is only one aspect of the photosynthetic characteristics of a leaf.
  The obtained results must be discussed in relation to the spectral distribution of measuring light so as not to make biased evaluations.

------

**Abbreviations**
BR-light: blue and red LED light,
Chl: chlorophyll,
ETR: electron transport rate,
FR: far-red,
LED: light-emitting diode,
*P*~n~: net photosynthetic rate,
PFD: photon flux density,
PPFD: photosynthetic photon flux density,
PSI: photosystem I,
PSII: photosystem II,

------

```{r, child = "./Main.Rmd"}
```

# References